---
title: "Representation Bias in Probability and Non-Probability Surveys - Theoretical Considerations and Practical Applications Using the sampcompR R-Package"
format: html
editor: visual
---

# Prepare for Exercises:

First we need to install some needed packages. The most important here is sampcompR, as it plays a prominent role in this course.

```{r Load backages}
library(sampcompR)
library(ggplot2)
library(xtable)
library(tidyverse)
library(mice)
```

Next we load data prepared for the excercieses:

```{r Load data}
load("generated_surveys/survey1_prob.RData")
load("generated_surveys/survey2_nonprob.RData")
load("generated_surveys/unobserved1_prob.RData")
load("generated_surveys/unobserved2_nonprob.RData")
load("generated_surveys/unobserved1_prob.RData")
load("generated_surveys/unobserved2_nonprob.RData")
load("generated_surveys/frame1_prob.RData")
load("generated_surveys/frame2_nonprob.RData") # normaly this would not exist

```

# Exercises

## Exercise 1: Response Model

Think about a probability Survey (Survey 1): Which X variables can you use from the frame? Which Y variables do you plan to analyze? Run a model, logistic regression model (nonresponse model) to investigate which variables led to nonresponse and by what extent. Interpret the findings. What do you learn from it?

```{r Exercise 1}

## Ex 1: nonresponse model
svy1.response <- glm(response ~ , 
                     family = "binomial", data = frame1)
summary(svy1.response)



```

## Exercise 2: R-Indicator

Use the R_indicator function of sampcompR to compute the R-indicator for your survey using frame Information.

What happens to the R-indicator if you drop an inﬂuential X-variable from the nonresponse model?

```{r, Exercise 2}

## if you need a hint try help(R_indicator)

## R indicator based on all frame variables
R_indicator(dfs=,
            variables=,
            response_identificators = ,
            get_r2 = T)

## R indicator when some influential variables are not used


```

## Exercise 3: Bias in univariate estimates

Install sampcompR from CRAN ( install.packages(sampcompR) )

Use sampcompR to estimate bias in univariate estimates for Survey 1 and Survey 2 (or for your own data if you like).

Plot the results and output them as a table.

How would you interpret the results?

```{r, Exercise 3}

### load a named vector of benchmark means
load("generated_surveys/benchmark.RData")
benchmark <- colMeans(mikrozensus)

### let us also assume that the full_probability survey with 
### no missings is our benchmark survey

full_prob<-cbind(frame1,unobserved1_prob[,-1])
benchmark_svy<-full_prob

###################################################
## benchmark available as named vector of means: ##
###################################################

## Generate Results
diff_pop_mic <-  uni_compare(nboots = 0,
                             funct= ,
                             dfs = ,
                             benchmarks=,
                             summetric = "avg2",
                             variables=)


## Plot Results
plot_diff_pop_mic<-plot_uni_compare(diff_pop_mic)
plot_diff_pop_mic

## Show Results in a Table
uni_diff_table<-uni_compare_table()
uni_diff_table


```

## Exercise 4 Bias in bivariate estimates

Use sampcompR to estimate bias in bivariate estimates for survey 1 and survey 2 (or for your own data if you like).

Plot the results.

How would you interpret the results?

```{r Exercise 4}


biv.diff <- biv_compare(nboots = , 
                        dfs = ,
                        benchmarks=,
                        diff_perc_size = 2,
                        plots_label = ,
                        variables=)

plot_biv_compare(biv.diff)




```

## Exercise 5 Bias in multivariate estimates

Use sampcompR to estimate bias in multivariate estimates for Survey 1 and Survey 2 (or for your own data if you like).

Plot the results.

How would you interpret he results?

```{r Exercise 5}


## multivariate comparison for dataset 1
multi.diff.1 <-  multi_compare(df=, 
                               bench=,
                               independent = c(),
                               dependent=c(),
                               family="ols")

## multivariate comparison for dataset 2
multi.diff.2 <- multi_compare(df=, 
                              bench=,
                              independent = c() ,
                              dependent=c(),
                              family="ols")

## Plot multivariate comparison
plot_multi_compare(c("multi.diff.1", "multi.diff.2"), diff_perc_size = 3, ,plots_label = c("survey 1", "survey 2")) + theme(legend.position="none")




```

## Exercise 6: Nonresponse weights

Use the response models from Exercise 1 to calculate nonresponse weights as inverse of the fitted values.

Investigate the weights regarding their summary and standard deviation.

Use them in sampcompR to univestigate their impact on univariate bias. (take a look in the documentation)

```{r, Exercise 6}





```

## Exercise 7: Correlation of y and nonresponse weights

Determine the correlation of your nonresponse weights and your variables.

Are there differences between your individual models?

```{r Exercise 7}




```

## Excursus: Raking weights for nonprobability surveys

We prepared an example of how you can include raking weights directly calculated by sampcompR in your survey comparison.

Take some time and try it out.

Interpret the results.

```{r Excursus}


### As we have only a dataframe with frame we also need to add information
### that would normaly not be observed

full_prob<-cbind(frame1,unobserved1_prob[,-1])
benchmark_svy<-full_prob


## a function to prepare weighting targets
targetlist<-function (target,df,weight=NULL,strata=NULL,id=NULL) {
  
  if(is.null(weight)==F) weight<-df[,weight]
  if(is.null(id)==F) id<-df[,id]
  if(is.null(id)==T) id<-c(1:nrow(df))
  if(is.null(strata)==F) strata<-df[,strata]
  target<-paste0("~",target)
  
  design <- survey::svydesign(ids = id, weights = weight, strata = strata, data=df)
  
  prop.table(survey::svytable(as.formula(target),design))
}


target_vars<-c("age_16to29", "age_30to39" ,"age_40to49",
               "age_50to59", "age_60p", "german" ,"female")



## prepare weighting targets
targets_svy<-map(target_vars,~targetlist(.,benchmark_svy, 
                                         id = NULL, 
                                         weight = NULL,
                                         strata = NULL))
names(targets_svy)<-target_vars


target_list<-list(targets_svy, targets_svy)

target_vars_list<-list(target_vars,target_vars)

## estimate biases when raking is used
diff_pop_rake <-  uni_compare(nboots = 0,funct= "rel_prop", 
                              plot_title = "comparison of survey 1 respondents and nonrespondents", 
                              dfs = c("prob_resp", "nonprob_part"), 
                              summetric = "avg2",
                              benchmarks=c("benchmark_svy"), 
                              variables=c("edu_low","edu_mid","edu_high",
                                          "satisfaction", "itknowledge", 
                                          "itliteracy","usage"),
                              adjustment_weighting = "raking",
                              adjustment_vars = target_vars_list,
                              raking_targets = target_list)

plot_diff_pop_rake<-plot_uni_compare(diff_pop_rake)
plot_diff_pop_rake


diff_pop <-  uni_compare(nboots = 0,funct= "rel_prop", 
                              plot_title = "comparison of survey 1 respondents and nonrespondents", 
                              dfs = c("prob_resp", "nonprob_part"), 
                              summetric = "avg2",
                              benchmarks=c("benchmark_svy"), 
                              variables=c("edu_low","edu_mid","edu_high",
                                          "satisfaction", "itknowledge", 
                                          "itliteracy","usage"))

plot_diff_pop<-plot_uni_compare(diff_pop)
plot_diff_pop

```
